GVN-hoist: hoisting expressions from control flow

### 2016-08-04 hoist.tex
calls

\subsection{Reduce register pressure}
Following example explains how code hoisting can actually reduce the register pressure.
Consider the following example where the labels prefixed with 'p' represent the position of
instruction in a basic block (names prefixed with bb).

bb0: p0: b = ...
bb1: p1: c = ...
bb2: p2: if (c) jump bb3 else jump bb4
bb3: p3: a0 = b + c
bb4: p4: a1 = b + c
bb5: p5: d = phi {a0(bb3), a1(bb4)}

@@

old-live-range = max(distance(p0, p3) + distance(p1, p3) , distance(p0, p4) + distance(p1, p4)) + distance (p3, p5)
new-live-rance = distance(p0, p2) + distance(p1, p2) + distance (p2, p5)

If the new live-range is less than the old one it will be a good candidate
for hoist. There are cases when both the ranges will be same:
\begin{enumerate}
  - There is only one operand on the right hand side e.g., an assignment operation.
  - One of the operands is not a Instruction/Register.
\end{enumerate}

@@

Hoisting scalars across calls is tricky because it can increase the
number of spills.

\subsection{In the presence of calls}
Caller saved registers.

@@

\section{Experimental Evaluation}

@@

\subsection{Optimistic approach}
In this approach, the goal is to maximise the total number of hoistings in the
entire function.  This algorithm is very useful when optimizing for code-size.
We see a lot of computations hoisted \ref{experimental-results}. In order to do
that, we collect the GVN of all the instructions in the function. After that we
iterate on the list of instructions having identical GVN. After that we find the
common dominator dominating all the identical computations and try to hoist the
instruction in that basic block, if feasible. Many times it is not possible to
hoist all the instructions to one common dominator, due to legality/profitability
constraints. In those cases, the algorithmn would partion the list of computations
a partially hoist a set of computations to their common dominator. For more details
see \ref{sec:partition}

@@

\subsection{Partition the list of hoisting candidates to maximize hoisting}
\cite{sec:partition}
The set of candidates are sorted in the order of their DFS-in
numbers. So that candidates close to each other in the list are also
closer (??) in the CFG.

>>>>

\subsection{Reduce register pressure}
\label{hoist:reg-pressure}
Following example explains how code hoisting can actually reduce the register pressure.
Consider the following example where the labels prefixed with 'p' represent the position of
instruction in a basic block (names prefixed with bb).

\begin{verbatim}
bb0: p0: b = ...
bb1: p1: c = ...
bb2: p2: if (c) jump bb3 else jump bb4
bb3: p3: a0 = b + c
bb4: p4: a1 = b + c
bb5: p5: d = phi {a0(bb3), a1(bb4)}

@@

old-live-range = max(distance(p0, p3) + distance(p1, p3) ,
                     distance(p0, p4) + distance(p1, p4))
                 + distance (p3, p5)
new-live-rance = distance(p0, p2) + distance(p1, p2)
                 + distance (p2, p5)
\end{verbatim}

If the new live-range is less than the old one it will be a good candidate
for hoist. There are cases when both the ranges will be same:
\begin{enumerate}
  \item There is only one operand on the right hand side e.g., an assignment operation.
  \item One of the operands is not a Instruction/Register.
\end{enumerate}

@@

\subsection{In the presence of calls}
Hoisting scalars across calls is tricky because it can increase the number of
spills. During the frame lowering of calls, the argument registers, in general,
the caller saved registers are saved because they might be modified by the
callee and after the call they are restored. So before the call the register
pressure is high because the number of available registers are reduced by the
number of caller saved registers. In that situation if a computation is hoisted
across the call, that would increase the total number of registers required by
one, thus contributing to the register pressure.

However, in the two special cases discussed in Section-\ref{hoist:reg-pressure},
it will be okay to hoist.

@@

\section{Experimental Evaluation}
\label{sec:experimental-results}

@@

\subsection{Optimistic approach}
In this approach, the goal is to maximise the total number of hoistings in the
entire function.  This algorithm is very useful when optimizing for code-size.
We see a lot of computations hoisted (see
Section~\ref{sec:experimental-results}. In order to do that, we collect the GVN
of all the instructions in the function. After that we iterate on the list of
instructions having identical GVN. After that we find the common dominator
dominating all the identical computations and try to hoist the instruction in
that basic block, if feasible. Many times it is not possible to hoist all the
instructions to one common dominator, due to legality/profitability
constraints. In those cases, the algorithmn would partion the list of
computations a partially hoist a set of computations to their common
dominator. For more details see Section~\ref{subsec:partition}.

@@

\subsection{Partition the list of hoisting candidates to maximize hoisting}
\label{subsec:partition}
The set of candidates are sorted in the order of their DFS-in
numbers. So that candidates close to each other in the list are also
closer (??) in the CFG.
